Playing with PCAs

### 2024-03-15 3-reports/playing_with_pcas.Rmd
do_pca function (for grouping)
pca by site

@@

>>>>

## BY SITE
```{r}

pca_owc = fit_pca_function(data_wide_PCA %>% filter(site == "OWC")) 
pca_ptr = fit_pca_function(data_wide_PCA %>% filter(site == "PTR")) 
pca_crc = fit_pca_function(data_wide_PCA %>% filter(site == "CRC")) 
pca_gcw = fit_pca_function(data_wide_PCA %>% filter(site == "GCW")) 
pca_msm = fit_pca_function(data_wide_PCA %>% filter(site == "MSM")) 
pca_gwi = fit_pca_function(data_wide_PCA %>% filter(site == "GWI"))

gg_pca_owc = ggbiplot(pca_owc$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_owc$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(color = groups))+ 
  labs(shape="", title = "PCA OWC", subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_pca_ptr = ggbiplot(pca_ptr$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_ptr$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(color = groups))+ 
  labs(shape="", title = "PCA PTR", subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_pca_crc = ggbiplot(pca_crc$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_crc$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(color = groups))+ 
  labs(shape="", title = "PCA CRC", subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_pca_gcw = ggbiplot(pca_gcw$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_gcw$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(color = groups))+ 
  labs(shape="", title = "PCA GCW", subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_pca_msm = ggbiplot(pca_msm$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_msm$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(color = groups))+ 
  labs(shape="", title = "PCA MSM", subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

gg_pca_gwi = ggbiplot(pca_gwi$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_gwi$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(color = groups))+ 
  labs(shape="", title = "PCA GWI", subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

library(patchwork)

gg_pca_gcw + gg_pca_msm + gg_pca_gwi +
  plot_layout(guides = "collect")&
  theme(legend.position = "bottom")

gg_pca_owc + gg_pca_ptr + gg_pca_crc +
  plot_layout(guides = "collect")&
  theme(legend.position = "bottom")

@@

```{r}
do_pca = function(dat){

  x = pca$pca_int$rotation %>% as.data.frame() %>% rownames_to_column("variable")

x = data_wide_PCA %>% group_by(region, transect) %>% do(do_pca(.))

x %>% 
  ggplot(aes(x = PC1, y = PC2, label = variable))+
  geom_point()+geom_text()+
  facet_grid(transect ~ region)
